---
title: "ROHs and IBD analyses (barplots)"
output: html_notebook
---
1. Libraries
```{r}
library(tidyr)
library(ggplot2)
library(sjmisc)
library(magrittr)
library(igraph)
```



2. Load files
```{r}
#set directory
datapath = '~\\masterBIOINFO\\practiques_tesis\\practiques\\data'
setwd(datapath)

## ROHS FILES###
#load individual files & snp files
indfile = 'mergeLAZ_FM_PJLYRICEU_comsnp_ld_roh.hom.indiv'
snpfile = 'mergeLAZ_FM_PJLYRICEU_comsnp_ld_roh.hom'
individuals = read.table(indfile, header= T)
snproh = read.table(snpfile, header= T)
#change size of roh's lenghts
individuals$mb = individuals$KB/1000
snproh$mb = snproh$KB/1000

###IBDS FILES###
#ibdfile = 'IBDsubset_ibd.ibd'
ibdfile = 'iberianpeninsula_allsnp_ids_ibd.ibd'
ibdtable= read.table(ibdfile, header= F)

ibdtable$len = ibdtable$V7 -ibdtable$V6
ibdtable$lenmb = ibdtable$len / 1000000

```

3. Define colours (palette)
```{r}
#blue
blue = rgb(128, 156, 200, maxColorValue = 255)
#pink
pink = rgb(208, 146, 166, maxColorValue = 255)
#yellow
yell = rgb(236, 214, 052, maxColorValue = 255)
#orange
oran = rgb(244, 164, 072, maxColorValue = 255)
#purple
purp = rgb(175, 129, 192, maxColorValue = 255)
#green
gree = rgb(120, 208, 104, maxColorValue = 255)

dblue = rgb(032, 058, 123, maxColorValue = 255)
dgree= rgb(026, 139, 017, maxColorValue = 255)
#targed red
redd = rgb(216, 000, 047, maxColorValue = 255)

```

###Analyses for rohs by individual ###

4.1 Group populations (verbose)

```{r}
spain =c("Basque", "Spanish", "Spanish_North", "Canary_Islander")
midleeast = c("Armenian", "Assyrian","Iranian", "Egyptian","Abkhasian", "Georgian","Jordanian", "Lebanese","Iranian_Bandari","Lebanese_Christian","Lebanese_Muslim","Druze", "Palestinian", "BedouinA", "BedouinB","Saudi","Syrian","Turkish","Jew_Turkish","Jew_Yemenite",  "Yemen","Jew_Georgian", "Jew_Iranian", "Jew_iraqi")
europe = c("Albanian","Greek","Bulgarian", "Croatian","Estonian","Lithuanian","Ukrainian",  "Hungarian","Belarusian","Romanian","Norwegian","Icelandic","Finnish", "Saami_WGA" ,"Italian_South", "Cypriot", "Czech","Sicilian",  "Maltese","Italian_North", "English", "Scottish" , "Orcadian","Basque", "Spanish", "Canary_Islander","Spanish_North","French","Jew_Ashkenazi", "FAG", "Russian")
cauc = c( "Adygei", "Aleut",     "Aleut_Tlingit",  "Altaian","Balkar", "Chechen",   "Chukchi",   "Chuvash",  "Dolgan",   "Eskimo","Even",  "Ignore_Adygei(relative_of_HGDP01382)" ,"Itelmen", "Kalmyk","Koryak",    "Kumyk","Lezgin",   "Mansi", "Mordovian","Nganasan",  "Nogai","North_Ossetian", "Selkup", "Tuvinian","Ulchi", "Yakut","Tubalar", "Yukagir")
mer = c("MER")
sar = c("Sardinian")
yri = c("YRI")
pjl= c("PJL")
ceu = c("CEU")
africa = c("Algerian","Jew_Libyan","Jew_Moroccan","Jew_Tunisian","Libyan","Moroccan","Mozabite","Tunisian")


pops1 = list(cauc,midleeast, europe,    sar, mer, pjl, ceu, yri, africa, spain)
popscolour = c(pink, oran, blue, blue, redd, yell,blue, gree, gree, purp)
#popsoutine = c(pink, oran, blue, blue, redd, yell, "black", "black", gree)
```
classify sub pops (pch)
```{r}
basque = c("Basque" , "Spanish_North" ) #square
jew = c("Jew_Ashkenazi","Jew_Georgian", "Jew_Iranian", "Jew_iraqi", "Jew_Libyan", "Jew_Moroccan", "Jew_Tunisian", "Jew_Turkish", "Jew_Yemenite") #triangle
roma = c('FAG', 'PJL') #PCH 25 inverted trianglet
target = c('MER')
refer = c('CEU', 'YRI')
#sumary + assgign colour


pops2 = list( target, '', basque,jew,  roma, refer)
pops2pch = c(19, 1,0,2,6, 10)
pchlabels = c("Mercheros", "", "basque", "jew",  "roma", "refer")

```


5. add labels
```{r}
#column for population
individuals$colour="grey"
#individuals$outcolour="grey"

for (i in 1:length(pops1)){
  subpop = pops1[[i]]
  for (j in subpop){
    individuals$colour[individuals$FID==j]=popscolour[i]
  }  
}

#column for shape
individuals$shp = 1
#pch
for (i in 1:length(pops2)){
  subpop = pops2[[i]]
  #print(subpop)
  #print(pops2pch[i])
  for (j in subpop){
    #print(j)
    individuals$shp[individuals$FID==j]=pops2pch[i]
  }
}
#p <- as_factor(individuals$shp)
```


6. PLOT ROHs by individual
```{r}
#kb - lentgh of all ROH
#nseg - number of ROH segments , (nroh)
labbb = c("Europe", "Basque", "Spanish Roma", "CEU", "Ashkenazi Jews", "Nord Africa", "African Jews", "Yoruba", "Middle East","Middle East Jews","Punjabi", "Rusia", "Mercheros" )
pchs = c(1, 0,6,10,2,1,2,10,1,2,6,1,19)

#nolegends
ggplot(data=individuals, aes(x=mb,y=NSEG, color=colour,shape=shp)) +
  geom_point() +
  labs( x = "Total length in ROH (Mb)",  y= "Number of ROH")+
  scale_color_identity(name = "location",
                       breaks= colours,
                       labels =  labbb)+
  scale_continuous_identity(name = "location",
                       #c("mer", "b", "basque","d", "e", "ref"),
                       #c(19, 1, 0, 2, 6, 10),
                       breaks = pchs,
                       labels =  labbb,
                       aesthetics = 'shape') +

  geom_point(data=individuals[individuals$FID=='MER',], aes(x=mb,y=NSEG) )

```


```{r}
#subgroup 
popssubg3 = c("CEU" ,"MER","YRI")
individuals4 =individuals[individuals$FID %in% popssubg3,]
ggplot(data=individuals4, aes(x=mb,y=NSEG, color=colour,shape=shp)) +
  geom_point() +
  labs( x = "Total length in ROH (Mb)",  y= "Number of ROH")+
  scale_color_identity(name = "location",
                       breaks= c(redd, blue,purp,  pink, dblue,gree,  dgree, yell, oran ),) +
  scale_shape_identity(name = "location",
                       breaks = as.character(c(19, 1,1,1,  1, 1,  1,    1,     1)),) +
  scale_fill_identity(guide = guide_legend())+
  geom_point(data=individuals[individuals$FID=='MER',], aes(x=mb,y=NSEG) )
```

###Analysis for the barplots##

1. Group them
```{r}
p_ib= c("Spanish")
basque = c("Basque", "Spanish_North")
sard = c("Sardinian")
mer = c("MER")
yri = c("YRI")
pjl= c("PJL")
roma = c("FAG")
ceu = c("CEU")
mozabite = c("Mozabite")

p_ib2 = c("Spanish_Castilla_y_Leon_IBS" , "Spanish_Castilla_la_Mancha_IBS" ,"Spanish_Extremadura_IBS", "Spanish_Cantabria_IBS" , "Spanish_Cataluna_IBS","Spanish_Valencia_IBS" , "Spanish_Baleares_IBS" ,       "Spanish_Murcia_IBS","Spanish_Andalucia_IBS" ,  "Spanish_Aragon_IBS","Spanish_Galicia_IBS")
#p_ib2 = c("Spanish_Castilla_y_Leon_IBS" , "Spanish_Castilla_la_Mancha_IBS" ,"Spanish_Extremadura_IBS", "Spanish_Cantabria_IBS" , "Spanish_Cataluna_IBS","Spanish_Valencia_IBS" ,        "Spanish_Murcia_IBS","Spanish_Andalucia_IBS" ,  "Spanish_Aragon_IBS","Spanish_Galicia_IBS")
basque2 = c("Spanish_Pais_Vasco_IBS" ,"Basque_Spanish",'Basque_French' ) 
pops3=list(mer, p_ib2, basque2)
pops2name3 = c( 'Mercheros','Spain','Basques')
pops2colour3 = c(redd,blue, purp)

pops2 = list(mer, p_ib, basque, sard,ceu, mozabite, yri,  pjl,  roma)
pops2name = c( 'Mercheros','Spain','Basques', 'Sardinia','Northern Europeans from Utah','Mozabite','Youruba','Punjabi',   'Roma' )
pops2colour = c(redd,blue, purp, pink,dblue, gree, dgree , yell,   oran)


#ibdtable
ibdtable=separate(ibdtable, V1, c("pop", "id"), sep="_(?=[^_]+$)")
ibdtable=separate(ibdtable, V3, c("pop2", "id2"), sep="_(?=[^_]+$)")
subibdtable = ibdtable[ ibdtable$pop== ibdtable$pop2,] #save just comparisons within pop

```

2. apply filter for lenght = total lenghts by individuals
```{r}
mblim = c(0.5,1,2,4,8,16)
#groups length intervals
len_catg = c("<0.5","0.5-1","1-2","2-4","4-8", "8-16", ">16")

#assign the value
snproh$len_group="<0.5"
for (i in 1:length(mblim)){
  snproh$len_group[snproh$mb > mblim[i]] = len_catg[i+1]
}
subibdtable$len_group="<0.5"
for (i in 1:length(subibdtable)){
  subibdtable$len_group[subibdtable$lenmb > mblim[i]] = len_catg[i+1]
}

len_catg = len_catg[2:7]


#create new dataset with snp data mean by individual
IID <-unique(snproh$IID)
ind_rohs <-data.frame(IID)

#for each group create a column
for(i in 1:length(len_catg)){
  ind_rohs[,len_catg[i]]=0
}

ind_rohs$pop="NOT ASSIGNED" #avoid the annoying na
ind_rohsnum = ind_rohs


#assign the total roh total lenght (by indivivual and toh group lenght)
for (i in IID){
  for(j in 1:length(len_catg)){
  rohs = snproh$mb[snproh$IID==i & snproh$len_group==len_catg[j]]
  ind_rohs[,len_catg[j]][ind_rohs$IID==i] = sum(rohs)
  ind_rohsnum[,len_catg[j]][ind_rohsnum$IID==i] = length(rohs)
  }
  ind_rohs$FID[ind_rohs$IID==i]=as.character(snproh$FID[snproh$IID==i])[1]
  ind_rohsnum$FID[ind_rohsnum$IID==i]=as.character(snproh$FID[snproh$IID==i])[1]
}

#assign the location group
for (i in 1:length(pops2)){
  subpop = pops2[[i]]
  for (j in subpop){
    ind_rohs$pop[ind_rohs$FID==j] = pops2name[i]
    ind_rohsnum$pop[ind_rohsnum$FID==j] = pops2name[i]
  }  
}




```

3. get populations means  + stdev
```{r}
#create summary dataset
rohlen=c()
pop= c()
for (i in 1:length(pops2name)){
  for (j in 1:length(len_catg)){
    pop= c(pop, pops2name[i])
    rohlen= c(rohlen, len_catg[j])
  }
}
barplot_table <- data.frame(rohlen,pop)
#add mean and stdev columns
barplot_table$mean = 0
barplot_table$stdev= 0
barplot_table$nummean = 0
barplot_table$numstdev= 0
barplot_table$repr= 0

#assgin values
for (i in 1:length(pops2name)){
  pop = pops2name[i]
  for (j in 1:length(len_catg)){
    #rohs length
    listpop=ind_rohs[,len_catg[j]][ind_rohs$pop==pop]
    barplot_table$mean[barplot_table$pop== pop & barplot_table$rohlen == len_catg[j]] = mean(listpop)
    barplot_table$stdev[barplot_table$pop== pop & barplot_table$rohlen == len_catg[j]] = sd(listpop)
    #rohs num
    listpop=ind_rohsnum[,len_catg[j]][ind_rohsnum$pop==pop]
    barplot_table$nummean[barplot_table$pop== pop & barplot_table$rohlen == len_catg[j]] = mean(listpop)
    barplot_table$numstdev[barplot_table$pop== pop & barplot_table$rohlen == len_catg[j]] = sd(listpop)
    #% individuals
    barplot_table$repr[barplot_table$pop== pop & barplot_table$rohlen == len_catg[j]] = 100 * length(listpop[listpop!=0]) /(length(listpop))
  }
}

```

5. plot
```{r}
barplot_table$rohlen <- factor(barplot_table$rohlen, levels = len_catg)
barplot_table$pop <- factor(barplot_table$pop, levels = pops2name)


ggplot(barplot_table, aes(x=as.factor(rohlen), y=mean, fill=pop)) +
  geom_bar(position=position_dodge(), stat="identity", colour='black') +
  geom_errorbar(aes(ymin=mean-stdev, ymax=mean+stdev), width=.2,position=position_dodge(.9))+
  labs( x = "ROH length category (Mb)",  y= "mean total ROH length (Mb)")+
  scale_fill_manual(values=pops2colour, labels=pops2name)+
  theme(legend.position="bottom")

ggplot(barplot_table, aes(x=as.factor(rohlen), y=nummean, fill=pop)) +
  geom_bar(position=position_dodge(), stat="identity", colour='black') +
  geom_errorbar(aes(ymin=nummean-numstdev, ymax=nummean+numstdev), width=.2,position=position_dodge(.9))+
  labs( x = "ROH length category (Mb)",  y= "mean total number of ROH")+
  scale_fill_manual(values=pops2colour, labels=pops2name)+
  theme(legend.position="bottom")

ggplot(barplot_table, aes(x=as.factor(rohlen), y=repr, fill=pop)) +
  geom_bar(position=position_dodge(), stat="identity", colour='black') +
  labs( x = " ROH length category (Mb)",  y= "% of individuals")+
  scale_fill_manual(values=pops2colour, labels=pops2name)+
  theme(legend.position="bottom")

setwd(datapath)
filetosave = paste(datapath,'\\roh_all.csv', sep='')
write.table(barplot_table, file = filetosave, sep = "\t", col.names = TRUE)

```
6. Repeat without the 3 outliers
```{r}
outliers = c ("MER2", "MER14", "MER11")
for (j in outliers){
  ind_rohs$pop[ind_rohs$IID==j]="NOT ASSIGNED"
  ind_rohsnum$pop[ind_rohsnum$IID==j]="NOT ASSIGNED"
}


#create summary dataset
rohlen=c()
pop= c()
for (i in 1:length(pops2name)){
  for (j in 1:length(len_catg)){
    pop= c(pop, pops2name[i])
    rohlen= c(rohlen, len_catg[j])
  }
}
barplot_table <- data.frame(rohlen,pop)
#add mean and stdev columns
barplot_table$mean = 0
barplot_table$stdev= 0
barplot_table$nummean = 0
barplot_table$numstdev= 0
barplot_table$repr= 0

#assgin values
for (i in 1:length(pops2name)){
  pop = pops2name[i]
  for (j in 1:length(len_catg)){
    #rohs length
    listpop=ind_rohs[,len_catg[j]][ind_rohs$pop==pop]
    barplot_table$mean[barplot_table$pop== pop & barplot_table$rohlen == len_catg[j]] = mean(listpop)
    barplot_table$stdev[barplot_table$pop== pop & barplot_table$rohlen == len_catg[j]] = sd(listpop)
    #rohs num
    listpop=ind_rohsnum[,len_catg[j]][ind_rohsnum$pop==pop]
    barplot_table$nummean[barplot_table$pop== pop & barplot_table$rohlen == len_catg[j]] = mean(listpop)
    barplot_table$numstdev[barplot_table$pop== pop & barplot_table$rohlen == len_catg[j]] = sd(listpop)
    #% individuals
    barplot_table$repr[barplot_table$pop== pop & barplot_table$rohlen == len_catg[j]] = 100 * length(listpop[listpop!=0]) /(length(listpop))
  }
}


```

6.2 plot
```{r}
barplot_table$rohlen <- factor(barplot_table$rohlen, levels = len_catg)
barplot_table$pop <- factor(barplot_table$pop, levels = pops2name)

ggplot(barplot_table, aes(x=as.factor(rohlen), y=mean, fill=pop)) +
  geom_bar(position=position_dodge(), stat="identity", colour='black') +
  geom_errorbar(aes(ymin=mean-stdev, ymax=mean+stdev), width=.2,position=position_dodge(.9))+
  labs( x = "ROH length category (Mb)",  y= "mean total ROH length (Mb)")+
  scale_fill_manual(values=pops2colour, labels=pops2name)+
  theme(legend.position="bottom")

ggplot(barplot_table, aes(x=as.factor(rohlen), y=nummean, fill=pop)) +
  geom_bar(position=position_dodge(), stat="identity", colour='black') +
  geom_errorbar(aes(ymin=nummean-numstdev, ymax=nummean+numstdev), width=.2,position=position_dodge(.9))+
  labs( x = "ROH length category (Mb)",  y= "mean total number of ROH")+
  scale_fill_manual(values=pops2colour, labels=pops2name)+
  theme(legend.position="bottom")

ggplot(barplot_table, aes(x=as.factor(rohlen), y=repr, fill=pop)) +
  geom_bar(position=position_dodge(), stat="identity", colour='black') +
  labs( x = " ROH length category (Mb)",  y= "% of individuals")+
  scale_fill_manual(values=pops2colour, labels=pops2name)+
  theme(legend.position="bottom")

setwd(datapath)
filetosave = paste(datapath,'\\roh_sub.csv', sep='')
write.table(barplot_table, file = filetosave, sep = "\t", col.names = TRUE)

```

###IBD analyses

8. open file  and group
```{r}
ibdtable2=separate(ibdtable, id, c("pop", "id"), sep="_(?=[^_]+$)")
 #save just comparisons within pop

ibdtable$p1='UNASSIGNED'
ibdtable$p2='UNASSIGNED'

#assign the location group (group them)
for (i in 1:length(pops3)){
  subpop = pops3[[i]]
  for (j in subpop){
    ibdtable$p1[ibdtable$pop==j] = pops2name3[i]
    ibdtable$p2[ibdtable$pop2==j] = pops2name3[i]
  }  
}
#save the pairs of IBD within the same population
subibdtable = ibdtable[ ibdtable$p1== ibdtable$p2,]

#assign lentgh group
len_catg = c("<0.5","0.5-1","1-2","2-4","4-8", "8-16", ">16")
#assign the value
subibdtable$len_group="<0.5"
for (i in 1:length(subibdtable)){
  subibdtable$len_group[subibdtable$lenmb > mblim[i]] = len_catg[i+1]
}
len_catg = len_catg[2:7]


#check for all individuals in the subset
IID = unique(subibdtable$id)
IID = c(IID, unique(subibdtable$id2))
IID = unique (IID)

#Create new tables: ibd len + ibd num
ind_ibd <-data.frame(IID)

#for each IND create a column
for(i in 1:length(len_catg)){
  ind_ibd[,len_catg[i]]=0
}

ind_ibd$pop='unass'
ind_ibdnum = ind_ibd


#assign the total ibd total lenght (by group and toh group lenght) to each individual - add pop!
for (i in IID){
  for(j in 1:length(len_catg)){
  ibds = subibdtable$lenmb[subibdtable$id==i & subibdtable$len_group==len_catg[j]]
  ibds2 = subibdtable$lenmb[subibdtable$id2==i & subibdtable$len_group==len_catg[j]]
  ibds = c(ibds, ibds2)
  ind_ibd[,len_catg[j]][ind_ibd$IID==i] = sum(ibds)
  ind_ibdnum[,len_catg[j]][ind_ibdnum$IID==i] = length(ibds)
  }
  pop = c(subibdtable$p1[subibdtable$id==i][1],subibdtable$p1[subibdtable$id2==i][1])
  pop=na.omit(pop)[1]
  ind_ibd$pop[ind_ibd$IID==i]=pop
  ind_ibdnum$pop[ind_ibdnum$IID==i]=pop
}



ibdlen=c()
pop= c()
for (i in 1:length(pops2name3)){
  for (j in 1:length(len_catg)){
    pop= c(pop, pops2name3[i])
    ibdlen= c(ibdlen, len_catg[j])
  }
}

ibdbarplot_table <- data.frame(ibdlen,pop)
#add mean and stdev columns
ibdbarplot_table$mean = 0
ibdbarplot_table$stdev= 0
ibdbarplot_table$nummean = 0
ibdbarplot_table$numstdev= 0
ibdbarplot_table$repr= 0

#assgin values
for (i in 1:length(pops2name3)){
  pop = pops2name[i]
  #print(pop)
  for (j in 1:length(len_catg)){
    #rohs length
    listpop=ind_ibd[,len_catg[j]][ind_ibd$pop==pop]
    ibdbarplot_table$mean[ibdbarplot_table$pop== pop & ibdbarplot_table$ibdlen == len_catg[j]] = mean(listpop)
    ibdbarplot_table$stdev[ibdbarplot_table$pop== pop & ibdbarplot_table$ibdlen == len_catg[j]] = sd(listpop)
    #rohs num
    listpop=ind_ibdnum[,len_catg[j]][ind_ibdnum$pop==pop]
    ibdbarplot_table$nummean[ibdbarplot_table$pop== pop & ibdbarplot_table$ibdlen == len_catg[j]] = mean(listpop)
    ibdbarplot_table$numstdev[ibdbarplot_table$pop== pop & ibdbarplot_table$ibdlen == len_catg[j]] = sd(listpop)
    #% individuals
    ibdbarplot_table$repr[ibdbarplot_table$pop== pop & ibdbarplot_table$ibdlen == len_catg[j]] = 100 * length(listpop[listpop!=0]) /(length(listpop))
  }
}

```
```{r}
#PLOT

ibdbarplot_table$ibdlen <- factor(ibdbarplot_table$ibdlen, levels = len_catg)
ibdbarplot_table$pop <- factor(ibdbarplot_table$pop, levels = pops2name3)


ggplot(ibdbarplot_table, aes(x=as.factor(ibdlen), y=mean, fill=pop)) +
  geom_bar(position=position_dodge(), stat="identity", colour='black') +
  geom_errorbar(aes(ymin=mean-stdev, ymax=mean+stdev), width=.2,position=position_dodge(.9))+
  labs( x = "IBD length category (Mb)",  y= "mean total IBD length (Mb)")+
  scale_fill_manual(values=pops2colour3, labels=pops2name3)+
  theme(legend.position="bottom")

ggplot(ibdbarplot_table, aes(x=as.factor(ibdlen), y=nummean, fill=pop)) +
  geom_bar(position=position_dodge(), stat="identity", colour='black') +
  geom_errorbar(aes(ymin=nummean-numstdev, ymax=nummean+numstdev), width=.2,position=position_dodge(.9))+
  labs( x = "IBD length category (Mb)",  y= "mean total number of IBD")+
  scale_fill_manual(values=pops2colour3, labels=pops2name3)+
  theme(legend.position="bottom")

ggplot(ibdbarplot_table, aes(x=as.factor(ibdlen), y=repr, fill=pop)) +
  geom_bar(position=position_dodge(), stat="identity", colour='black') +
  labs( x = "IBD length category (Mb)",  y= "% of individuals")+
  scale_fill_manual(values=pops2colour3, labels=pops2name3)+
  theme(legend.position="bottom")

```

9. save data
```{r}
setwd(datapath)
filetosave = paste(datapath,'\\ibd_pib_all.csv', sep='')
write.table(ibdbarplot_table, file = filetosave, sep = "\t", col.names = TRUE)

```
10. repeat without outliers
```{r}
#remove ROH outliers

for (j in outliers){
  ind_ibd$pop[ind_ibd$IID==j]="NOT ASSIGNED"
  ind_ibdnum$pop[ind_ibdnum$IID==j]="NOT ASSIGNED"
}


ibdlen=c()
pop= c()
for (i in 1:length(pops2name3)){
  for (j in 1:length(len_catg)){
    pop= c(pop, pops2name3[i])
    ibdlen= c(ibdlen, len_catg[j])
  }
}

ibdbarplot_table <- data.frame(ibdlen,pop)
#add mean and stdev columns
ibdbarplot_table$mean = 0
ibdbarplot_table$stdev= 0
ibdbarplot_table$nummean = 0
ibdbarplot_table$numstdev= 0
ibdbarplot_table$repr= 0

#assgin values
for (i in 1:length(pops2name3)){
  pop = pops2name[i]
  #print(pop)
  for (j in 1:length(len_catg)){
    #rohs length
    listpop=ind_ibd[,len_catg[j]][ind_ibd$pop==pop]
    ibdbarplot_table$mean[ibdbarplot_table$pop== pop & ibdbarplot_table$ibdlen == len_catg[j]] = mean(listpop)
    ibdbarplot_table$stdev[ibdbarplot_table$pop== pop & ibdbarplot_table$ibdlen == len_catg[j]] = sd(listpop)
    #rohs num
    listpop=ind_ibdnum[,len_catg[j]][ind_ibdnum$pop==pop]
    ibdbarplot_table$nummean[ibdbarplot_table$pop== pop & ibdbarplot_table$ibdlen == len_catg[j]] = mean(listpop)
    ibdbarplot_table$numstdev[ibdbarplot_table$pop== pop & ibdbarplot_table$ibdlen == len_catg[j]] = sd(listpop)
    #% individuals
    ibdbarplot_table$repr[ibdbarplot_table$pop== pop & ibdbarplot_table$ibdlen == len_catg[j]] = 100 * length(listpop[listpop!=0]) /(length(listpop))
  }
}


setwd(datapath)
filetosave = paste(datapath,'\\ibd_pib_sub.csv', sep='')
write.table(ibdbarplot_table, file = filetosave, sep = "\t", col.names = TRUE)

```


```{r}
#PLOT

ibdbarplot_table$ibdlen <- factor(ibdbarplot_table$ibdlen, levels = len_catg)
ibdbarplot_table$pop <- factor(ibdbarplot_table$pop, levels = pops2name3)


ggplot(ibdbarplot_table, aes(x=as.factor(ibdlen), y=mean, fill=pop)) +
  geom_bar(position=position_dodge(), stat="identity", colour='black') +
  geom_errorbar(aes(ymin=mean-stdev, ymax=mean+stdev), width=.2,position=position_dodge(.9))+
  labs( x = "IBD length category (Mb)",  y= "mean total IBD length (Mb)")+
  scale_fill_manual(values=pops2colour3, labels=pops2name3)+
  theme(legend.position="bottom")

ggplot(ibdbarplot_table, aes(x=as.factor(ibdlen), y=nummean, fill=pop)) +
  geom_bar(position=position_dodge(), stat="identity", colour='black') +
  geom_errorbar(aes(ymin=nummean-numstdev, ymax=nummean+numstdev), width=.2,position=position_dodge(.9))+
  labs( x = "IBD length category (Mb)",  y= "mean total number of IBD")+
  scale_fill_manual(values=pops2colour3, labels=pops2name3)+
  theme(legend.position="bottom")

ggplot(ibdbarplot_table, aes(x=as.factor(ibdlen), y=repr, fill=pop)) +
  geom_bar(position=position_dodge(), stat="identity", colour='black') +
  labs( x = "IBD length category (Mb)",  y= "% of individuals")+
  scale_fill_manual(values=pops2colour3, labels=pops2name3)+
  theme(legend.position="bottom")

```
11.compare (files were merged in excel)
```{r}
setwd(datapath)
ibdvsroh = 'rohsvsibd.csv'
ibdrohtable= read.table(ibdvsroh, header= T, sep=';')
groups = unique(ibdrohtable$pop)
groups = c("Mercheros ROHs", "Mercheros ROHs subset", "Mercheros", "Mercheros subset","Spain ROHs","Spain","Spain peninsula","Basques ROHs", "Basques")

ibdrohtable2=ibdrohtable[ibdrohtable$ibdlen != '0.5-1',]

ibdrohtable = ibdrohtable2[ibdrohtable2$pop != "Spain peninsula",]

ibdrohtable$ibdlen <- factor(ibdrohtable$ibdlen, levels = len_catg[2:6])
ibdrohtable$pop <- factor(ibdrohtable$pop, levels = groups)
groups = c("Mercheros ROHs", "Mercheros ROHs subset", "Mercheros IBD", "Mercheros IBD subset","Spain ROHs","Spain IBD ","Spain peninsula IBD","Basques ROHs", "Basques IBD")
colour3 = c(redd, pink, oran, yell, dblue, blue, purp, dgree, gree)
groups = c("Mercheros ROHs", "Mercheros ROHs subset", "Mercheros IBD", "Mercheros IBD subset","Spain ROHs","Spain IBD ","Basques ROHs", "Basques IBD")
colour3 = c(redd, pink, oran, yell, dblue, blue, dgree, gree)


ggplot(ibdrohtable, aes(x=as.factor(ibdlen), y=mean, fill=pop)) +
  geom_bar(position=position_dodge(), stat="identity", colour='black') +
  geom_errorbar(aes(ymin=mean-stdev, ymax=mean+stdev), width=.2,position=position_dodge(.9))+
  labs( x = "length category (Mb)",  y= "mean total length (Mb)")+
  scale_fill_manual(values=colour3, labels=groups)+
  theme(legend.position="bottom")

ggplot(ibdrohtable, aes(x=as.factor(ibdlen), y=nummean, fill=pop)) +
  geom_bar(position=position_dodge(), stat="identity", colour='black') +
  geom_errorbar(aes(ymin=nummean-numstdev, ymax=nummean+numstdev), width=.2,position=position_dodge(.9))+
  labs( x = "length category (Mb)",  y= "mean total number")+
  scale_fill_manual(values=colour3, labels=groups)+
  theme(legend.position="bottom")

ggplot(ibdrohtable, aes(x=as.factor(ibdlen), y=repr, fill=pop)) +
  geom_bar(position=position_dodge(), stat="identity", colour='black') +
  labs( x = "length category (Mb)",  y= "% of individuals")+
  scale_fill_manual(values=colour3, labels=groups)+
  theme(legend.position="bottom")

```